Assessing primitives performance\\ on multi-stage execution

### 2018-05-18 paper.tex
update plot and table
update figures in paragraphs

2   23.492410   0.0134896734
3   22.541872   0.077485955
4   24.927379   0.03087542

@@

   \end{subfigure}
   \subcaption*{\textbf{\LLVMMac}\vspace{-0.2cm}} %change vspace based on fig position
    \end{subfigure}
   \caption{String comparison benchmarks for strings of different sizes on different architectures. Average relative speed-up regarding baseline. Higher than 1 is better.}
   \label{fig:bench}
\end{figure*}
\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
    & & 0 & 1 & 5 & 100 & 1000 \\
\hline
  \multirow{2}{*}{\legSlangPlug} &\GCCLinux& 0.413 $\pm$ 0.0005 & 0.746 $\pm$ 0.0008 & 1.804 $\pm$ 0.0005 & 12.047 $\pm$ 0.0088 & 23.492 $\pm$ 0.0135 \\
  						&\LLVMMac& 0.339 $\pm$ 0.0007 & 0.6238 $\pm$ 0.0012 & 1.5807 $\pm$ 0.0092 & 11.4096 $\pm$ 0.0474  & 23.21 $\pm$ 0.1789 \\
   \hline
  \multirow{2}{*}{\legSlang} &\GCCLinux& 1.287 $\pm$ 0.0048 & 2.257 $\pm$ 0.0116 & 4.958 $\pm$ 0.0111 & 17.241 $\pm$ 0.0515 & 22.542 $\pm$ 0.0775\\
 					 &\LLVMMac& 0.932 $\pm$ 0.0015 & 1.6900 $\pm$ 0.0030 & 3.241 $\pm$ 0.0067 & 10.149 $\pm$ 0.0321  & 12.939 $\pm$ 0.0177 \\
   \hline
  \multirow{2}{*}{\legRTL}  &\GCCLinux& 3.343 $\pm$ 0.0211 & 5.171 $\pm$ 0.0265  & 8.715 $\pm$ 0.0221 & 17.657 $\pm$ 0.0481 & 24.927 $\pm$ 0.0309\\
 				 	&\LLVMMac & 3.255 $\pm$ 0.0079 & 5.216 $\pm$ 0.0153 & 8.971 $\pm$ 0.0359 & 17.643 $\pm$ 0.06  & 24.881 $\pm$ 0.1983 \\
  \hline
\end{tabular} 
\vspace{0.2cm}
\caption{Benchmark results with standard errors. Average relative speed-up regarding baseline.}
\label{tab:detail_bench}
\end{table*}

@@

\paragraph{Stack switch overhead.}
As mentioned in Section~\ref{sec:VMexec}, all implementations written in Slang come with an overhead due to the switch between the C and Smalltalk stacks. This overhead is clearly visible in the bench results in Figure~\ref{fig:bench} for small strings (size 1 or less): the Plugin implementation is between 57,8 (\GCCLinux) to 61,4\% (\LLVMMac) slower than the baseline's for empty strings.
As the string size goes larger, the overhead is less significant because it is absorbed by other performance gains.
Regarding this criteria, \legRTL version is always the fastest as it is not impacted by the switch overhead.

@@

\paragraph{Control over generated code.} One could notice a rather striking performance difference between the 2 compilers: when compiling with GCC, the performance between the \legSlang and \legRTL implementations is really close for long strings (100 characters and longer), with respectively 22.54 and 24.93 relative speed-up factors for a 1000-characters string. Yet performance greatly differs when LLVM is used, with respectively 12,94 and 24,88 relative speed-up factors.

>>>>

2   25.35945   0.1742098
3   25.42575   0.1621165
4   26.76261   0.186075

@@

   \end{subfigure}
   \subcaption*{\textbf{\LLVMMac}\vspace{-0.2cm}} %change vspace based on fig position
    \end{subfigure}
   \caption{String comparison benchmarks for strings of different sizes on different architectures. Average relative speed-up regarding baseline. Higher than 1 is better.}
   \label{fig:bench}
\end{figure*}
\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
    & & 0 & 1 & 5 & 100 & 1000 \\
\hline
  \multirow{2}{*}{\legSlangPlug} &\GCCLinux& 0.413 $\pm$ 0.0005 & 0.746 $\pm$ 0.0008 & 1.804 $\pm$ 0.0005 & 12.047 $\pm$ 0.0088 & 25.359 $\pm$ 0.1742 \\
  						&\LLVMMac& 0.339 $\pm$ 0.0007 & 0.6238 $\pm$ 0.0012 & 1.5807 $\pm$ 0.0092 & 11.4096 $\pm$ 0.0474  & 23.21 $\pm$ 0.1789 \\
   \hline
  \multirow{2}{*}{\legSlang} &\GCCLinux& 1.287 $\pm$ 0.0048 & 2.257 $\pm$ 0.0116 & 4.958 $\pm$ 0.0111 & 17.241 $\pm$ 0.0515 & 25.426 $\pm$ 0.1621\\
 					 &\LLVMMac& 0.932 $\pm$ 0.0015 & 1.6900 $\pm$ 0.0030 & 3.241 $\pm$ 0.0067 & 10.149 $\pm$ 0.0321  & 12.939 $\pm$ 0.0177 \\
   \hline
  \multirow{2}{*}{\legRTL}  &\GCCLinux& 3.343 $\pm$ 0.0211 & 5.171 $\pm$ 0.0265  & 8.715 $\pm$ 0.0221 & 17.657 $\pm$ 0.0481 & 26.763 $\pm$ 0.1861\\
 				 	&\LLVMMac & 3.255 $\pm$ 0.0079 & 5.216 $\pm$ 0.0153 & 8.971 $\pm$ 0.0359 & 17.643 $\pm$ 0.06  & 24.881 $\pm$ 0.1983 \\
  \hline
\end{tabular} 
\vspace{0.2cm}
\caption{Benchmark results with standard errors. Average relative speed-up regarding baseline.}
\label{tab:detail_bench}
\end{table*}

@@

\paragraph{Stack switch overhead.}
As mentioned in Section~\ref{sec:VMexec}, all implementations written in Slang come with an overhead due to the switch between the C and Smalltalk stacks. This overhead is clearly visible in the bench results in Figure~\ref{fig:bench} for small strings (size 1 or less): the Plugin implementation is between 58,7 (\GCCLinux) to 66,1\% (\LLVMMac) slower than the baseline's for empty strings.
As the string size goes larger, the overhead is less significant because it is absorbed by other performance gains.
Regarding this criteria, \legRTL version is always the fastest as it is not impacted by the switch overhead.

@@

\paragraph{Control over generated code.} One could notice a rather striking performance difference between the 2 compilers: when compiling with GCC, the performance between the \legSlang and \legRTL implementations is really close for long strings (100 characters and longer), with respectively 25.24 and 26.76 relative speed-up factors for a 1000-characters string. Yet performance greatly differs when LLVM is used, with respectively 12,94 and 24,88 relative speed-up factors.
